---
title: "capstone graphics"
author: "Elena Flauto -- 505697812"
date: "2025-02-27"
output: html_document
---


spotify monthly listeners vs superfans
```{r}
artist <- c("Rihanna", "Taylor Swift", "Drake", "Post Malone", "Ed Sheeran", "Travis Scott", "Kanye West", "Lana Del Rey", "Future", "Arctic Monkeys")
artist_pop_rank <- c(89, 96, 94, 87, 86, 92, 91, 89, 90, 86)
proportion_spotify <- c(0.5825123, 0.3007220, 0.1887383, 0.3560732, 0.2743191, 0.4069675, 0.2828467, 0.2898551, 0.1045752, 0.6523810)
spotify_monthly_listeners <- c(89433621, 86845825, 74236296,69545833, 69129305, 71669551, 66102111, 59837435, 58077087, 55040913)
subscribers_count <- c(44100000, 60600000, 30500000, 27900000, 55800000, 19500000, 11400000, 16800000, 14200000, 8670000)
instagram_followers <- c(1.50e+08, 2.82e+08, 1.43e+08, 2.70e+07, 4.84e+07, 5.88e+07, 2.08e+07, 2.04e+07, 2.63e+07, 7.70e+06)
engagement_rates <- c(1.358048, 2.210638, 1.083175, 2.217683, 1.141767, 3.870708, 4.698937, 15.166550, 2.874610, 7.941482)
live_score <- c(0.0, 6.0, 2.7, 1.8, 6.0, 2.7, 3.6, 1.8, 2.0, 2.0)
full_data_table <- cbind(artist, artist_pop_rank, proportion_spotify, spotify_monthly_listeners, subscribers_count, instagram_followers, engagement_rates, live_score)
library(tidyverse)
tibble_data <- as_tibble(full_data_table)
print(tibble_data)

library(tidyr)
clean_data <- drop_na(tibble_data)
clean_data


clean_data <- clean_data %>%
  mutate(monthly_listeners = c(89433621, 86845825, 74236296, 69545833, 69129305, 71669551, 66102111, 59837435, 58077087, 55040913)) #taken feb 5 2025 

clean_data <- clean_data %>%
  mutate(instagram_followers = c(150000000, 282000000, 143000000, 27000000, 48400000, 58800000, 20800000, 20400000, 26300000, 7700000))

clean_data

#spotify_2024_rank <- c(21, 1, 3, 25, 36, 8, 6, 12, NA, NA)
#10 of the top 50 streamed artists in 2024

#superfan formula


clean_data <- clean_data %>%
  mutate(across(c(instagram_followers, subscribers_count, engagement_rates, live_score, artist_pop_rank, proportion_spotify), as.numeric))

clean_data <- clean_data %>%
  mutate(live_multiplier = (1+(live_score/6)))


#superfans_lb <- subscribers_count + instagram_followers * aligned_rates/100
clean_data <- clean_data %>%
  mutate(superfans_lb = monthly_listeners * .01)



clean_data <- clean_data %>%
  mutate(superfans_est = superfans_lb + ((instagram_followers * (engagement_rates / 100) + subscribers_count*(engagement_rates / 100)) * live_multiplier * proportion_spotify))

clean_data <- clean_data %>%
  mutate(superfans_ub = monthly_listeners * .1)

clean_data

# Load necessary libraries
library(ggplot2)
library(dplyr)

min_superfans <- min(clean_data$superfans_lb, clean_data$superfans_est, clean_data$superfans_ub, na.rm = TRUE)
max_superfans <- max(clean_data$superfans_lb, clean_data$superfans_est, clean_data$superfans_ub, na.rm = TRUE)

# Assuming clean_data_2 is your tibble

# Create the plot
ggplot(clean_data) +
  # Add bar graph for artist_pop_rank
  geom_bar(aes(x = artist, y = monthly_listeners), stat = "identity", fill = "#1ED760", alpha = 0.7) +
  
  # Add line graph for superfans_lb
  geom_line(aes(x = artist, y = superfans_lb, group = 1), color = "#000000", size = 1) +
  
  # Add line graph for superfans_est
  geom_line(aes(x = artist, y = superfans_est, group = 1), color = "purple", size = 1) +
  
  # Add line graph for superfans_ub
  geom_line(aes(x = artist, y = superfans_ub, group = 1), color = "#ff66c4", size = 1) +

  # Create y-axis
  scale_y_continuous(
    name = "Monthly Listeners (Millions)",
    labels = scales::label_comma(scale = 1e-6)
  ) +
  
  # scale_y_continuous(
  #   name = "Monthly Listeners",
  #   labels = scales::label_comma(scale = 1e-6),
  #   limits = c(min_superfans, max_superfans * 1.1),  # Zoom in on the lines portion (with a small margin above max)
  #   breaks = seq(min_superfans, max_superfans, by = (max_superfans - min_superfans) / 5)  # Adjust break intervals for clarity
  # ) +
  
  # Customize plot appearance
  theme_minimal() +
  labs(
    title = "Spotify Monthly Listeners vs Superfans",
    x = "Artists",
    y = "Monthly Listeners (Millions)"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x axis labels for readability
    axis.title.x = element_text(margin = margin(t = 10)),  # Add margin to x-axis title
    axis.title.y = element_text(margin = margin(r = 10))   # Add margin to y-axis title
  )



ggplot(clean_data) +
  # Add bar graph for artist_pop_rank
  geom_bar(aes(x = artist, y = monthly_listeners), stat = "identity", fill = "#1ED760", alpha = 0.7) +
  
  # Add line graph for superfans_lb
  geom_line(aes(x = artist, y = superfans_lb, group = 1), color = "#000000", size = 1) +
  
  # Add line graph for superfans_est
  geom_line(aes(x = artist, y = superfans_est, group = 1), color = "#1ed760", size = 1) +
  
  # Add line graph for superfans_ub
  geom_line(aes(x = artist, y = superfans_ub, group = 1), color = "#ff66c4", size = 1) +
  
  # # Create dual y-axes
  # scale_y_continuous(
  #   name = "Monthly Listeners",
  #   labels = scales::label_comma(scale = 1e-6)
  # ) +
  
  scale_y_continuous(
    name = "Monthly Listeners (Millions)",
    labels = scales::label_comma(scale = 1e-6),
    limits = c(min_superfans, max_superfans * 1.1),  # Zoom in on the lines portion (with a small margin above max)
    breaks = seq(min_superfans, max_superfans, by = (max_superfans - min_superfans) / 5)  # Adjust break intervals for clarity
  ) +
  
  # Customize plot appearance
  theme_minimal() +
  labs(
    title = "Estimated Spotify Superfans",
    x = "Artists",
    y = "Monthly Listeners"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x axis labels for readability
    axis.title.x = element_text(margin = margin(t = 10)),  # Add margin to x-axis title
    axis.title.y = element_text(margin = margin(r = 10))   # Add margin to y-axis title
  )


```
```{r}
library(knitr)
library(tools)

parameters <- c(
  "Artist Popularity Rank", 
  "Proportion of Listeners on Spotify", 
  "Spotify Followers", 
  "Spotify Monthly Listeners", 
  "YouTube Usernames and Subscriber Counts", 
  "Chartmetric IDs", 
  "Annual Average Instagram Engagement Rate", 
  "Instagram Followers and URLs", 
  "Most Recent Tour Year", 
  "Most Recent Tour Size", 
  "Active Touring", 
  "Bonus", 
  "Top 100 Artists", 
  "Top 37 Spotify Artists"
)

sources <- c(
  "(Spotify)", 
  "(Kworb)", 
  "(Spotify)", 
  "(Spotify)", 
  "(YouTube)", 
  "(Chartmetric)", 
  "(Charmetric)", 
  "(Instagram)", 
  "(Pollstar)", 
  "(Determined Rank)", 
  "(Manual)", 
  "(Pollstar)",  
  "(Kworb)", 
  "(Spotify)"
)

information <- c(
  "Assigned by Spotify", 
  "Artists' Spotify Listeners divided by Overall", 
  "", 
  "", 
  "", 
  "", 
  "Average Instagram engagement rate for the artist in 2024", 
  "", 
  "The most recent year in which the artist went on tour", 
  "Assigned value (1-5) based on venue capacities, with 5 being the greatest", 
  "Binary value indicating whether the artist is still touring", 
  "Bonus for artists who had a top-grossing or top-attended major tour", 
  "Globally as of Jan 2025", 
  "Globally as of 2024"
)

methods <- c(
  "Spotify API", 
  "Web Scraping", 
  "Spotify API", 
  "Spotify API", 
  "YouTube API", 
  "Chartmetric API", 
  "Chartmetric API", 
  "Web Scraping", 
  "Pollstar", 
  "Pollstar", 
  "Manual", 
  "Pollstar", 
  "Web Scraping", 
  "Spotify"
)

parameter_data <- data.frame(
  Parameters = parameters,
  Source = sources,
  Information = information,
  Method = methods
)

# Create the pretty table with a title
kable(parameter_data, caption = "Data Collected")

```

tiers
coefficients very similar even after lasso, ridge, and cross validation
v1 - top streamed
feature selection:
disc_ts      27.114125 - ea
win_ts       19.009962 - ea
mix_ts       15.690093 - hq
ea_ts        12.803606
--
snk_ts       12.217301
hq_ts        12.051637
unr_ts       10.900175
merch_ts      7.425253

correlated:
disc, ea
ea, win
mix, hq

 Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
 0.1250  0.8438  1.5625  1.8602  2.4375  5.2500 

base: 1.5626
tier 1: disc ($1.49)
tier 2: disc, ea, win ($1.99)
tier 3: mix ($1.49)
tier 4: mix, hq ($1.99)
tier 5: all ($2.99)

v2 - fav
feature selection
win_fv       31.735732 - ea, unr
stem_fv      22.187635 - hq
disc_fv      21.519307 - ea
ea_fv        20.070625 - merch
---
unr_fv       14.368962
merch_fv     13.645455
hq_fv        12.878398
snk_fv        8.037627

correlated:
disc, ea
ea, win
mix, hq
unr, win
merch, ea

   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
 0.0000  0.5938  1.3750  1.6743  2.2812  6.1250 
 
 base: $1.375
 tier 1: win ($1.49)
 tier 2: win, ea, unr ($1.99)
 tier 3: mix ($1.49)
 tier 4: mix, hq ($1.99)
 tier 5: win, disc, ea, unr, merch ($2.99)
 tier 6: all above ($3.49)
 

```{r}
# Prices for Version 1 (V1)
v1_prices <- c(1.49, 1.99, 1.49, 1.99, 2.99)  # Prices for tiers 1, 2, 3, 4, 5 in V1

# Prices for Version 2 (V2)
v2_prices <- c(1.49, 1.99, 1.49, 1.99, 2.99, 3.49)  # Prices for tiers 1, 2, 3, 4, 5, 6 in V2

clean_data %>%
  summarise(total_superfans_lb = sum(superfans_lb, na.rm = TRUE))

clean_data %>%
  summarise(total_superfans_est = sum(superfans_est, na.rm = TRUE))

clean_data %>%
  summarise(total_superfans_ub = sum(superfans_ub, na.rm = TRUE))

superfan_levels <- c(6999180, 20712707, 69991798)
```
version 1
```{r}
revenue_matrix <- outer(v1_prices, superfan_levels, FUN = "*")

# Convert the matrix into a tibble for better readability
library(tibble)
revenue_table <- as_tibble(revenue_matrix, 
                           .name_repair = "unique") %>%
  setNames(c("revenue_lb", "revenue_est", "revenue_ub")) %>%
  mutate(v1_prices = 1:5) %>%
  select(v1_prices, everything())

# View the result
colnames(revenue_table) <- c("Price Tiers", "Revenue Lower Bound", "Estimated Revenue", "Revenue Upper Bound")
print(revenue_table)

```
version 2
```{r}
revenue_matrix_2 <- outer(v2_prices, superfan_levels, FUN = "*")

# Convert the matrix into a tibble for better readability
library(tibble)
revenue_table_2 <- as_tibble(revenue_matrix_2, 
                           .name_repair = "unique") %>%
  setNames(c("revenue_lb", "revenue_est", "revenue_ub")) %>%
  mutate(v2_prices = 1:6) %>%
  select(v2_prices, everything())

# View the result

colnames(revenue_table_2) <- c("Price Tiers", "Revenue Lower Bound", "Estimated Revenue", "Revenue Upper Bound")
print(revenue_table_2)
```



```{r}
library(knitr)

revenue_table$`Price Tiers` <- c("$1.49", "$1.99", "$1.49", "$1.99", "$2.99")
revenue_table$`Revenue Lower Bound` <- paste0("$", format(revenue_table$`Revenue Lower Bound`, big.mark = ","))
revenue_table$`Estimated Revenue` <- paste0("$", format(revenue_table$`Estimated Revenue`, big.mark = ","))
revenue_table$`Revenue Upper Bound` <- paste0("$", format(revenue_table$`Revenue Upper Bound`, big.mark = ","))

kable(revenue_table, align = "l", caption = "Revenue Projections - Top-Streamed")

revenue_table_2$`Price Tiers` <- c("$1.49", "$1.99", "$1.49", "$1.99", "$2.99", "$3.49")
revenue_table_2$`Revenue Lower Bound` <- paste0("$", format(revenue_table_2$`Revenue Lower Bound`, big.mark = ","))
revenue_table_2$`Estimated Revenue` <- paste0("$", format(revenue_table_2$`Estimated Revenue`, big.mark = ","))
revenue_table_2$`Revenue Upper Bound` <- paste0("$", format(revenue_table_2$`Revenue Upper Bound`, big.mark = ","))

kable(revenue_table_2, align = "l", caption = "Revenue Projections - Favorite Artists")
```
Ideal Tier Model - Top-Streamed
```{r}
kable(revenue_table[c(2, 5), ], align = "l", caption = "Proposed Tiers and Revenue Projections - Top-Streamed")
```
Ideal Tier Model - Favorite Artists
```{r}
kable(revenue_table_2[c(2, 5, 6), ], align = "l", caption = "Proposed Tiers and Revenue Projections - Favorite Artists")
```

Figure 4: Spotify Average Revenue per User by Year
Show average revenue per user 2015 to 2023 (business of apps)
Year	Average revenue per user (€)
2015	6.84
2016	6.2
2017	5.32
2018	4.89
2019	4.86
2020	4.41
2021	4.29
2022	4.54
2023	4.27
2024	4.62
** add observation - potential revenue per use 2025
```{r}
ARPU_2024 <- 4.80  # Example ARPU for 2024 in USD
existing_users_2024 <- 675000000  # Total users for 2024
new_users_contributing <- 20712707 *5  # Users contributing the new revenue
#monthly_revenue_new_users_usd <- ((4.80+1.99)*0.90*new_users_contributing + (4.80+2.99)*0.08*new_users_contributing + (4.80+3.49)*0.02*new_users_contributing) / new_users_contributing # Monthly revenue from new users (USD)
monthly_revenue_new_users_usd <- ((4.80+1.99)*0.75*new_users_contributing + (4.80+2.99)*0.2*new_users_contributing + (4.80+3.49)*0.05*new_users_contributing) / new_users_contributing # Monthly revenue from new users (USD)

# Calculate the remaining users
remaining_users_2025 <- existing_users_2024 - new_users_contributing

# Revenue from remaining users (using 2024 ARPU)
monthly_revenue_remaining_users <- remaining_users_2025 * ARPU_2024  # Monthly revenue from remaining users

# Calculate ARPU for 2025
ARPU_2025 <- 4.80*(remaining_users_2025/existing_users_2024) + (monthly_revenue_new_users_usd)*(new_users_contributing/existing_users_2024)

# Display ARPU for 2025
ARPU_2025 <- round(ARPU_2025, 2)
ARPU_2025

apple_music_users <- 88e6
amazon_music_users <- 82e6
youtube_music_users <- 80e6

# Calculate 1% of each user base
one_percent_apple <- 0.01 * apple_music_users
one_percent_amazon <- 0.01 * amazon_music_users
one_percent_youtube <- 0.01 * youtube_music_users
 
# Fan Tier 1 price and months in a year
fan_tier_price <- 1.99
months_per_year <- 12
revenue_apple <- one_percent_apple * fan_tier_price * months_per_year
revenue_amazon <- one_percent_amazon * fan_tier_price * months_per_year
revenue_youtube <- one_percent_youtube * fan_tier_price * months_per_year

total_additional_revenue <- revenue_apple + revenue_amazon + revenue_youtube
total_additional_revenue

```

```{r}
# Data for Spotify Average Revenue per User by Year in EUR
average_revenue_per_user <- data.frame(
  Year = c(2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024),
  Average_Revenue_per_User_Euro = c(6.84, 6.20, 5.32, 4.89, 4.86, 4.41, 4.29, 4.54, 4.27, 4.62)
)

# Assuming the exchange rate is 1 EUR = 1.10 USD
exchange_rate <- 1.04

# Converting from EUR to USD
average_revenue_per_user$Average_Revenue_per_User_Euro <- round(average_revenue_per_user$Average_Revenue_per_User_Euro * exchange_rate, 2)
average_revenue_per_user$Average_Revenue_per_User_Euro <- as.numeric(average_revenue_per_user$Average_Revenue_per_User_Euro)
colnames(average_revenue_per_user) <- c("Year", "ARPU")

library(ggplot2)
ggplot(average_revenue_per_user, aes(x = Year, y = ARPU)) +
  geom_point(color = "#1ED760", size = 3) +  # Adding red dots at each point
  geom_line(color = "#1ED760", size = 1) +  # Line graph with blue color and width of 1
  labs(
    title = "Average Revenue per User Over Time",
    x = "Year",
    y = "ARPU (USD)"
  ) +
  scale_x_continuous(
    breaks = average_revenue_per_user$Year,  # Show every year on the x-axis
    labels = as.character(average_revenue_per_user$Year)  # Customize the labels (you can modify this to show specific formats)
  ) +
  theme_minimal() +  # Minimal theme for better aesthetics
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the title
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels
  )

library(knitr)
average_revenue_per_user$Year <- as.factor(average_revenue_per_user$Year)
average_revenue_per_user$ARPU <- paste0("$", format(average_revenue_per_user$ARPU, big.mark = ","))
# kable(average_revenue_per_user, align = "l", caption = "Average Revenue per User (USD)")

#add observation - potential revenue per user 2025

```

```{r}
# WITHOUT GETTING ANY NEW USERS OR DOING ANYTHING ELSE
# REASON FOR LOW AMT IS REGIONAL PRICING + 
# DOESN'T ACC FOR FREE USERS MOVING FROM PREMIUM TO THIS (JUST PREMIUM TO THIS)
# DISCUSS W RESPECT TO REGION AND COUNTRY GRAPHS
# Data for Spotify Average Revenue per User by Year in EUR
average_revenue_per_user <- data.frame(
  Year = c(2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024),
  Average_Revenue_per_User_Euro = c(6.84, 6.20, 5.32, 4.89, 4.86, 4.41, 4.29, 4.54, 4.27, 4.62)
)

# Assuming the exchange rate is 1 EUR = 1.04 USD
exchange_rate <- 1.04

# Converting from EUR to USD
average_revenue_per_user$Average_Revenue_per_User_Euro <- round(average_revenue_per_user$Average_Revenue_per_User_Euro * exchange_rate, 2)
average_revenue_per_user$Average_Revenue_per_User_Euro <- as.numeric(average_revenue_per_user$Average_Revenue_per_User_Euro)
colnames(average_revenue_per_user) <- c("Year", "ARPU")

# Add 2025 ARPU 
average_revenue_per_user <- rbind(average_revenue_per_user, data.frame(Year = 2025, ARPU = ARPU_2025))
average_revenue_per_user$Color <- ifelse(average_revenue_per_user$Year == 2025, "#ff66c4", "#1ED760")

# Plot the data
library(ggplot2)
ggplot(average_revenue_per_user, aes(x = Year, y = ARPU)) +
  geom_line(color = "#1ED760", size = 1) +  # Line graph with green color and width of 1
  geom_point(aes(color = Color), size = 3) +  # Adding green dots at each point
  labs(
    title = "Average Revenue per User Over Time",
    x = "Year",
    y = "ARPU (USD)"
  ) +
  scale_x_continuous(
    breaks = average_revenue_per_user$Year,  # Show every year on the x-axis
    labels = as.character(average_revenue_per_user$Year)  # Customize the labels (you can modify this to show specific formats)
  ) +
  theme_minimal() +  # Minimal theme for better aesthetics
  scale_color_identity() + 
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the title
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels
  ) +
geom_text(
    data = subset(average_revenue_per_user, Year == 2025),  # Filter for the 2025 data point
    aes(label = paste0("$", format(ARPU, big.mark = ","))),
    vjust = -1,  # Adjust vertical position of the label
    hjust = 0.5,  # Adjust horizontal position of the label
    color = "#000000",  # Label color to match the point color
    size = 3  # Adjust label size
  )

# Create a table with the updated data (including 2025)
library(knitr)
average_revenue_per_user$ARPU <- paste0("$", format(average_revenue_per_user$ARPU, big.mark = ","))
# kable(average_revenue_per_user, align = "l", caption = "Average Revenue per User (USD)")


```
```{r}
#superfan tier projection
ARPU_2024 <- 4.80  # Example ARPU for 2024 in USD
existing_users_2024 <- 675000000  # Total users for 2024
new_users_contributing <- 20712707*5  # Users contributing the new revenue
#monthly_revenue_new_users_usd <- ((4.80+1.99)*0.90*new_users_contributing + (4.80+2.99)*0.08*new_users_contributing + (4.80+3.49)*0.02*new_users_contributing) / new_users_contributing # Monthly revenue from new users (USD)
monthly_revenue_new_users_usd <- ((4.80+1.99)*0.75*new_users_contributing + (4.80+2.99)*0.2*new_users_contributing + (4.80+3.49)*0.05*new_users_contributing) # Monthly revenue from new users (USD)
annual_revenue_new_users <- monthly_revenue_new_users_usd * 12

# Calculate the remaining users
remaining_users_2025 <- existing_users_2024 - new_users_contributing

# Revenue from remaining users (using 2024 ARPU)
monthly_revenue_remaining_users <- remaining_users_2025 * ARPU_2024  # Monthly revenue from remaining users
annual_revenue_remaining_users <- monthly_revenue_remaining_users* 12

total_projected_annual_revenue_2025 <- annual_revenue_new_users + 15620000000 # = 24400116497
total_projected_annual_revenue_2025
```
Figure 3: Spotify Total, Premium Subscriber, and Ad-supported Revenue by Year
** add point for superfan tier projection 2025: ($24400116497)
Spotify revenue 2016 to 2024 ($bn) -boa
Year	Revenue (€bn)
2016	2.94
2017	4.62
2018	5.25
2019	6.76
2020	7.88
2021	9.66
2022	11.72
2023	13.24
2024	15.62

Here is a table displaying the Spotify revenue through Ad-supported users by year:
demandsage
Year	Revenue
2024 (Up to Q3)	€1.32 billion + .5b Q4
2023	€1.681 billion
2022	€1.476 billion
2021	€1.208 billion
2020	€745 million
2019	€678 million
2018	€542 million
2017	€416 million
2016	€295 million
2015	€196 million

Here is a table displaying Spotify’s revenue through Premium subscribers by year:
Year	Revenue
2024 (Up to Q3)	€10.12 billion + 3.705 Q4
2023	€11.566 billion
2022	€10.25 billion
2021	€8.46 billion
2020	€7.13 billion
2019	€6.086 billion
2018	€4.717 billion
2017	€3.67 billion
2016	€2.657 billion
2015	€1.74 billion
2014	€982.7 million
2013	€677.89 million
2012	€374.6 million

```{r}
# Revenue data (in EUR)
total_revenue_usd <- c(2.94, 4.62, 5.25, 6.76, 7.88, 9.66, 11.72, 13.24, 15.62)  # Total revenue in EUR (2016-2024)
ad_supported_revenue_eur <- c(0.295, 0.416, 0.542, 0.678, 0.745, 1.208, 1.476, 1.681, 1.85)  # Ad-supported revenue in EUR (2016-2024)
premium_revenue_eur <- c(2.657, 3.67, 4.717, 6.086, 7.13, 8.46, 10.25, 11.566, 10.12+3.705)  # Premium revenue in EUR (2016-2024)

# Conversion rate from EUR to USD
exchange_rate <- 1.04

# Convert from EUR to USD
ad_supported_revenue_usd <- ad_supported_revenue_eur * exchange_rate
premium_revenue_usd <- premium_revenue_eur * exchange_rate

# Data frame for Spotify revenues
years <- 2016:2024
revenues <- data.frame(
  Year = years,
  Total_Revenue = total_revenue_usd,
  Ad_Supported_Revenue = ad_supported_revenue_usd,
  Premium_Revenue = premium_revenue_usd
)


# Plot the data
library(ggplot2)

ggplot(revenues, aes(x = Year)) +
  geom_line(aes(y = Total_Revenue, color = "Total Revenue"), size = 1) +
  geom_point(aes(y = Total_Revenue, color = "Total Revenue"), size = 3) +  # Add points to Total Revenue line
  geom_line(aes(y = Ad_Supported_Revenue, color = "Ad-Supported Revenue"), size = 1) +
  geom_point(aes(y = Ad_Supported_Revenue, color = "Ad-Supported Revenue"), size = 3) +  # Add points to Ad-Supported Revenue line
  geom_line(aes(y = Premium_Revenue, color = "Premium Revenue"), size = 1) +
  geom_point(aes(y = Premium_Revenue, color = "Premium Revenue"), size = 3) +  # Add points to Premium Revenue line
  labs(
    title = "Spotify Revenue by Type (in Billions USD)",
    x = "Year",
    y = "Revenue (USD Billions)"
  ) +
  scale_color_manual(values = c(
    "Total Revenue" = "#1ED760", 
    "Ad-Supported Revenue" = "#000000", 
    "Premium Revenue" = "purple", 
    "Superfan Tier" = "#ff66c4"  # Same color for Premium Revenue and Superfan Tier
  ), aesthetics = "color") +
  scale_x_continuous(
    breaks = seq(2016, 2025, by = 1),  # Show every year
    labels = as.character(2016:2025)  # Customize the labels
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  guides(color = guide_legend(title = "Revenue Type"))

# Print the data table
library(knitr)
revenues$Total_Revenue <- paste0("$", format(revenues$Total_Revenue, big.mark = ","))
revenues$Ad_Supported_Revenue <- paste0("$", format(revenues$Ad_Supported_Revenue, big.mark = ","))
revenues$Premium_Revenue <- paste0("$", format(revenues$Premium_Revenue, big.mark = ","))
# kable(revenues, align = "l", caption = "Spotify Revenue by Year (USD)")

```

```{r}
# Revenue data (in EUR)
total_revenue_usd <- c(2.94, 4.62, 5.25, 6.76, 7.88, 9.66, 11.72, 13.24, 15.62)  # Total revenue in EUR (2016-2024)
ad_supported_revenue_eur <- c(0.295, 0.416, 0.542, 0.678, 0.745, 1.208, 1.476, 1.681, 1.85)  # Ad-supported revenue in EUR (2016-2024)
premium_revenue_eur <- c(2.657, 3.67, 4.717, 6.086, 7.13, 8.46, 10.25, 11.566, 10.12+3.705)  # Premium revenue in EUR (2016-2024)

# Conversion rate from EUR to USD
exchange_rate <- 1.04

# Convert from EUR to USD
ad_supported_revenue_usd <- ad_supported_revenue_eur * exchange_rate
premium_revenue_usd <- premium_revenue_eur * exchange_rate

# Data frame for Spotify revenues
years <- 2016:2024
revenues <- data.frame(
  Year = years,
  Total_Revenue = total_revenue_usd,
  Ad_Supported_Revenue = ad_supported_revenue_usd,
  Premium_Revenue = premium_revenue_usd
)


# Add the Superfan Tier projection and Premium Revenue for 2025
revenues <- rbind(revenues, data.frame(
  Year = 2025,
  Total_Revenue = total_projected_annual_revenue_2025/1000000000,  # Already in billions
  Ad_Supported_Revenue = NA,  # Not provided for Superfan Tier
  Premium_Revenue = premium_revenue_usd[9] + (annual_revenue_new_users/1000000000) # Premium Revenue for 2025
))

# Plot the data
library(ggplot2)

ggplot(revenues, aes(x = Year)) +
  geom_line(aes(y = Total_Revenue, color = "Total Revenue"), size = 1) +
  geom_point(aes(y = Total_Revenue, color = "Total Revenue"), size = 3) +  # Add points to Total Revenue line
  geom_line(aes(y = Ad_Supported_Revenue, color = "Ad-Supported Revenue"), size = 1) +
  geom_point(aes(y = Ad_Supported_Revenue, color = "Ad-Supported Revenue"), size = 3) +  # Add points to Ad-Supported Revenue line
  geom_line(aes(y = Premium_Revenue, color = "Premium Revenue"), size = 1) +
  geom_point(aes(y = Premium_Revenue, color = "Premium Revenue"), size = 3) +  # Add points to Premium Revenue line
  geom_point(data = subset(revenues, Year == 2025), aes(y = Total_Revenue, color = "Superfan Tier"), size = 3) + 
  geom_point(data = subset(revenues, Year == 2025), aes(y = Premium_Revenue, color = "Superfan Tier"), size = 3) +  # Same color for Premium 2025
  labs(
    title = "Spotify Revenue by Type (in Billions USD)",
    x = "Year",
    y = "Revenue (USD Billions)",
    fill = "Revenue Type"
  ) +
  scale_color_manual(values = c(
    "Total Revenue" = "#1ED760", 
    "Ad-Supported Revenue" = "#000000", 
    "Premium Revenue" = "purple", 
    "Superfan Tier" = "#ff66c4"  
  )) +
  scale_x_continuous(
    breaks = seq(2016, 2025, by = 1),  # Show every year
    labels = as.character(2016:2025)  # Customize the labels
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  guides(color = guide_legend(title = "Revenue Type"))

# Print the data table
library(knitr)
revenues$Total_Revenue <- paste0("$", format(revenues$Total_Revenue, big.mark = ","))
revenues$Ad_Supported_Revenue <- paste0("$", format(revenues$Ad_Supported_Revenue, big.mark = ","))
revenues$Premium_Revenue <- paste0("$", format(revenues$Premium_Revenue, big.mark = ","))
# kable(revenues, align = "l", caption = "Spotify Revenue by Year (USD)")

```

figure 2: spotify users and subscribers by region - bo
Spotify users by region 2016 to 2023 (mm)
Year	Europe	North America	Latin America	Rest of World
2016	40	36	20	8
2017	51	46	28	13
2018	66	56	38	20
2019	83	65	49	35
2020	102	77	66	54
2021	124	88	80	73
2022	138	97	95	103
2023	159	111	116	165
2024	169	107	138	212

Spotify subscribers by region 2018 to 2024 (mm)
Year	Europe	North America	Latin America	Rest of World
2018	33	26	17	7
2019	43	32	22	11
2020	54	40	29	15
2021	66	48	33	18
2022	73	54	39	22
2023	86	62	46	26
2024	92	64	55	35

```{r}
# Data for Spotify Users by Region (2016-2024)
users_data <- data.frame(
  Year = 2016:2024,
  Europe = c(40, 51, 66, 83, 102, 124, 138, 159, 169),
  North_America = c(36, 46, 56, 65, 77, 88, 97, 111, 107),
  Latin_America = c(20, 28, 38, 49, 66, 80, 95, 116, 138),
  Rest_of_World = c(8, 13, 20, 35, 54, 73, 103, 165, 212)
)

# Data for Spotify Subscribers by Region (2018-2024)
subscribers_data <- data.frame(
  Year = 2018:2024,
  Europe = c(33, 43, 54, 66, 73, 86, 92),
  North_America = c(26, 32, 40, 48, 54, 62, 64),
  Latin_America = c(17, 22, 29, 33, 39, 46, 55),
  Rest_of_World = c(7, 11, 15, 18, 22, 26, 35)
)

colnames(users_data) <- c("Year", "Europe", "North America", "Latin America", "Rest of World")
colnames(subscribers_data) <- c("Year", "Europe", "North America", "Latin America", "Rest of World")

# Loading knitr library
library(knitr)

# Create the kable for Spotify Users by Region
kable(users_data, align = "l", caption = "Spotify Users by Region (Millions)")

# Create the kable for Spotify Subscribers by Region
kable(subscribers_data, align = "l", caption = "Spotify Subscribers by Region (Millions)")

```

```{r}
# Load necessary libraries
library(tidyr)
library(ggplot2)

# Data for Spotify Users by Region (2016-2024)
users_data <- data.frame(
  Year = 2016:2024,
  Europe = c(40, 51, 66, 83, 102, 124, 138, 159, 169),
  North_America = c(36, 46, 56, 65, 77, 88, 97, 111, 107),
  Latin_America = c(20, 28, 38, 49, 66, 80, 95, 116, 138),
  Rest_of_World = c(8, 13, 20, 35, 54, 73, 103, 165, 212)
)

# Data for Spotify Subscribers by Region (2018-2024)
subscribers_data <- data.frame(
  Year = 2018:2024,
  Europe = c(33, 43, 54, 66, 73, 86, 92),
  North_America = c(26, 32, 40, 48, 54, 62, 64),
  Latin_America = c(17, 22, 29, 33, 39, 46, 55),
  Rest_of_World = c(7, 11, 15, 18, 22, 26, 35)
)

# Reshape Users Data to Long Format
users_data_long <- pivot_longer(users_data, 
                                cols = -Year, 
                                names_to = "Region", 
                                values_to = "Users")

# Reshape Subscribers Data to Long Format
subscribers_data_long <- pivot_longer(subscribers_data, 
                                      cols = -Year, 
                                      names_to = "Region", 
                                      values_to = "Subscribers")

# Add "Type" column to differentiate Users vs Subscribers
users_data_long$Type <- "Users"
subscribers_data_long$Type <- "Subscribers"

# Debugging step: Check the column names to ensure they match before combining
print("Users Data Columns:")
print(colnames(users_data_long))
print("Subscribers Data Columns:")
print(colnames(subscribers_data_long))

# Ensure the column names are consistent
colnames(users_data_long) <- c("Year", "Region", "value", "Type")
colnames(subscribers_data_long) <- c("Year", "Region", "value", "Type")

# Combine Users and Subscribers Data
combined_data <- rbind(users_data_long, subscribers_data_long)

combined_data$Region <- gsub("_", " ", combined_data$Region)

# Check the first few rows of the combined data to ensure everything looks correct
print("Combined Data:")
head(combined_data)

# Plotting the Histogram
ggplot(combined_data, aes(x = Year, y = value, fill = Type)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.4) +  # Use dodge to separate the bars
  scale_fill_manual(values = c("Users" = "#1ED760", "Subscribers" = "purple")) +  # Custom colors for users and subscribers
  facet_wrap(~ Region, scales = "free_y", labeller = label_value) +  # Separate by region, allow different y-axis scales
  labs(title = "Spotify Users vs. Subscribers by Region (2016-2024)",
       x = "Year",
       y = "Millions",
       fill = "Type") +
  scale_x_continuous(
    breaks = 2016:2024,  # Show every year
    labels = as.character(2016:2024)
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for readability

```

figure 1: music users and user demographics by country - potentially unnecessary

Nation	Gen Z	Millennials	Gen X	Boomers
Brazil	87%	89%	73%	63%
Mexico	83%	87%	81%	75%
Colombia	84%	88%	76%	69%
Argentina	87%	89%	77%	58%
Chile	91%	92%	78%	72%
France	84%	68%	53%	37%
Spain	82%	78%	74%	57%
UK	77%	79%	67%	41%
Italy	77%	76%	59%	40%
Germany	83%	78%	59%	37%
Japan	89%	70%	51%	34%
US	99%	98%	96%	89%

Over half of Sweden’s population uses a paid music streaming service (IFPI)
According to IFPI, Sweden is the most engaged country in terms of paid music subscriptions with 61% of the population falling under this description.

Here are the top five nations:

Sweden – 61%
Mexico – 57%
Germany – 55%
USA – 53%
New Zealand – 52%
```{r}
# Create a new dataset for the top 5 countries with highest paid music streaming percentages
top5_countries <- data.frame(
  Nation = c("Sweden", "Mexico", "Germany", "USA", "New Zealand"),
  Percentage = c(61, 57, 55, 53, 52)
)

# Plot top 5 countries with custom hex colors for each country
ggplot(top5_countries, aes(x = reorder(Nation, -Percentage), y = Percentage, fill = "#1ed760")) +
  geom_bar(stat = "identity", fill = "#1ed760", col = NA) +
  labs(title = "Top 5 Countries with Paid Music Streaming Service Usage",
       x = "Country", y = "Percentage") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme_minimal()
```

```{r}
# Load necessary libraries
library(ggplot2)
library(reshape2)

# Create the dataset
data <- data.frame(
  Nation = c("Brazil", "Mexico", "Colombia", "Argentina", "Chile", "France", 
             "Spain", "UK", "Italy", "Germany", "Japan", "US"),
  Gen_Z = c(87, 83, 84, 87, 91, 84, 82, 77, 77, 83, 89, 99),
  Millennials = c(89, 87, 88, 89, 92, 68, 78, 79, 76, 78, 70, 98),
  Gen_X = c(73, 81, 76, 77, 78, 53, 74, 67, 59, 59, 51, 96),
  Boomers = c(63, 75, 69, 58, 72, 37, 57, 41, 40, 37, 34, 89)
)

# Melt the data for plotting
data_melted <- melt(data, id.vars = "Nation", variable.name = "Generation", value.name = "Percentage")
data_melted$Generation <- gsub("_", " ", data_melted$Generation)

# Plot histogram using ggplot with hex color codes
ggplot(data_melted, aes(x = Nation, y = Percentage, fill = Generation)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Percentage of Each Generation Using Paid Music Streaming Service",
       x = "Country", y = "Percentage") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),  # Tilt x-axis labels at 45 degrees
    plot.title = element_text(hjust = 0.5)) +
  scale_fill_manual(values = c("Gen Z" = "#1ed760", 
                               "Millennials" = "#ff66c4", 
                               "Gen X" = "purple", 
                               "Boomers" = "orange")) +  # Custom hex colors
  theme_minimal()

```

